---
title: "Task_2_Project1"
Group Members: "Aartee Simran Dhomeja"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
markdown::markdownToHTML
```

```{r}
#Libraries
library(fpc)
library(readr)
library(stringr)
library(factoextra)
library(plotly)
library(tidyverse)
library(NbClust)
library(philentropy)
library(clValid)
library(kohonen)
library(ClusterR)
```

```{r}
#Data Set
World_Indicators<- read.csv('D:/FDA/Project 1/World Indicators.csv')

```

```{r}
#First Analysis
#GDP & Infant Mortality Rate

#Clean the data to only include required columns
World_IndicatorsGI <- World_Indicators[c(6,10,21)]

#Remove NAs from the data set
World_IndicatorsGI <- World_IndicatorsGI %>% drop_na()


#Determine the optimal number of clusters using the silhouette method
fviz_nbclust(World_IndicatorsGI[,1:2], kmeans, method = "silhouette")

#Determine the optimal number of clusters using the elbow method
fviz_nbclust(World_IndicatorsGI[,1:2], kmeans, method = "wss")

#Find the kmeans of GDP and Infant Mortality
set.seed(55)
cluster.World_IndicatorsGI <- kmeans(World_IndicatorsGI[,1:2], 2, nstart = 20)
cluster.World_IndicatorsGI
cluster.World_IndicatorsGI$cluster

#Graph of the GDP & Infectant Mortality Clustered
cluster.World_IndicatorsGI$cluster <- as.factor(cluster.World_IndicatorsGI$cluster)

WIGI_Clustered <- plot_ly(World_IndicatorsGI, x = ~GDP, 
                          y = ~ Infant.Mortality.Rate, 
                          color = ~cluster.World_IndicatorsGI$cluster)

WIGI_Clustered <- WIGI_Clustered %>% add_markers()
WIGI_Clustered <- WIGI_Clustered %>% layout(scene = 
                                              list(xaxis = list(title = 'X1'),
                                                   yaxis = list(title = 'X2')))
WIGI_Clustered <- WIGI_Clustered %>% layout(title = 'GDP vs. Infant Mortality Clustered')

WIGI_Clustered # plot 


#Hierarchical Clustering
#Using Hierarchical analysis to plot the clusters

#To find the distance between each clusters
dm_Gdp_infant <- dist(World_IndicatorsGI[,1:2], method = 'euclidean')

# Hierarchical Clustering

set.seed(50)
hc_Gdp_infant <- hclust(dm_Gdp_infant,method = "average")

#Plotting Dendogram for hierarchical cluster

hc_Gdp_infant_dendogram <- plot(hc_Gdp_infant, main = "GDP Vs Infant.Mortality.Rate Dendogram")

#Using Cut-tree to identify the number of clusters

abline(h = 3.3, col = "blue")
fit_Gdp_infant <- cutree(hc_Gdp_infant, k = 2)
rect.hclust(hc_Gdp_infant, k = 2, border = "red")

#Binding the data
World_IndicatorsGI <- cbind(World_IndicatorsGI, fit_Gdp_infant)

#Plotting hierarchical clusters

World_IndicatorsGI_hc_plot <- plot_ly(World_IndicatorsGI, x = ~GDP, y = ~Infant.Mortality.Rate, color = ~as.factor(World_IndicatorsGI$fit_Gdp_infant))
World_IndicatorsGI_hc_plot <- World_IndicatorsGI_hc_plot %>% add_markers()
World_IndicatorsGI_hc_plot <- World_IndicatorsGI_hc_plot %>% layout(title = 'GDP Vs Infant.Mortality.Rate hierarchical Clustered')
World_IndicatorsGI_hc_plot

#Self Organizing Maps - 

#Create the model
World_IndicatorsGI_SOM <- as.matrix(scale(World_IndicatorsGI[,1:2]))
World_IndicatorsGI_Grid <- somgrid(xdim = 5, ydim = 5, topo = "rectangular")
set.seed(55)
World_IndicatorsGI_SOM_Model <- som(X = World_IndicatorsGI_SOM, grid = World_IndicatorsGI_Grid)

#Plot the model
plot(World_IndicatorsGI_SOM_Model, type = "counts")

plot(World_IndicatorsGI_SOM_Model, type = "mapping")

plot(World_IndicatorsGI_SOM_Model, type = "codes")


#ADD INTERNAL VALIDATION 
#validation for 2 clusters
calinhara(World_IndicatorsGI[,1:2],cluster.World_IndicatorsGI$cluster,cn=2)

#validation for 3
cluster.World_IndicatorsGI_3 <- kmeans(World_IndicatorsGI[,1:2], 3, nstart = 20)
cluster.World_IndicatorsGI_3
calinhara(World_IndicatorsGI[,1:2],cluster.World_IndicatorsGI_3$cluster,cn=3)

#validation for 4
cluster.World_IndicatorsGI_4 <- kmeans(World_IndicatorsGI[,1:2], 4, nstart = 20)
cluster.World_IndicatorsGI_4
calinhara(World_IndicatorsGI[,1:2],cluster.World_IndicatorsGI_4$cluster,cn=4)

##################### Post CH Value Plot ##########################
WIGI_Clustered_CH <- plot_ly(World_IndicatorsGI, x = ~GDP, 
                             y = ~Infant.Mortality.Rate, 
                             color = ~cluster.World_IndicatorsGI_4$cluster)
WIGI_Clustered_CH <- WIGI_Clustered_CH %>% add_markers()
WIGI_Clustered_CH <- WIGI_Clustered_CH %>% layout(scene = 
                                                    list(xaxis = list(title = 'X1'),
                                                         yaxis = list(title = 'X2')))
WIGI_Clustered_CH <- WIGI_Clustered_CH %>% layout(
  title = 'GDP vs. Infant Mortality Clustered')

WIGI_Clustered_CH # plot 



#add dunn index and ch value

unique(World_IndicatorsGI$Clust)
typeof(cluster.World_IndicatorsGI$cluster)
cluster.World_IndicatorsGI$cluster<-as.numeric(
  cluster.World_IndicatorsGI$cluster)
cluster.stats(dist(World_IndicatorsGI[,1:2],method ='euclidean'),
              cluster.World_IndicatorsGI$cluster)

#cluster.World_IndicatorsGI$cluster<-as.numeric(cluster.World_IndicatorsGI$cluster)

x=dunn(clusters = cluster.World_IndicatorsGI$cluster, 
       Data = World_IndicatorsGI[,1:2])
cat("The value for dunn index is ,",x,"")

#Combine clusters with original table
World_IndicatorsGI$Clust <- cluster.World_IndicatorsGI$cluster
World_IndicatorsGI$Clust_CH <- cluster.World_IndicatorsGI_4$cluster
World_IndicatorsGI_Sub1<-subset(World_IndicatorsGI,Clust_CH==1)
World_IndicatorsGI_Sub2<-subset(World_IndicatorsGI,Clust_CH==2)
World_IndicatorsGI_Sub3<-subset(World_IndicatorsGI,Clust_CH==3)
World_IndicatorsGI_Sub4<-subset(World_IndicatorsGI,Clust_CH==4)
print(World_IndicatorsGI_Sub1,World_IndicatorsGI_Sub2,World_IndicatorsGI_Sub3,World_IndicatorsGI_Sub4)
#World_IndicatorsGI_Sub1

#China, Japan, and USA represent cluster 1, all other countries (excluding NA) are represented by cluster 2.


```
# the first analysis focuses on the link between GDP of a country and its infant mortality rate. The optimal number of clusters according to KMeans and Silhoutte is 2 but according to the CH Value, 4 is the ideal number of clusters. 
The Dunn Index and external validation has always been carried out. After clustering and plotting the same it could be analyzed that countries with minimal to low GDP have high infant mortality as compared to those with higher amount of gdp. For Example countries like fall in the second 2nd cluster for mortality

```{r}
#Second Analysis
#GDP vs HealthCapita

#Data
#World_Indicators <- read.csv("/Users/kailash/Downloads/World Indicators (2).csv")

##########Scaling ######################

World_Indicators_GH<- World_Indicators %>% select(GDP,Health.Exp.Capita,
                                                  Country)
World_Indicators_GH<-na.omit(World_Indicators_GH)
#class(World_Indicators_GH$Health.Exp.Capita)

World_Indicators_GH$Health.Exp.Capita <- as.numeric(World_Indicators_GH$Health.Exp.Capita)

World_Indicators_GH <- World_Indicators_GH %>% drop_na()

World_Indicators_GHS<-as.data.frame(scale(World_Indicators_GH[,1:2]))
#DF<-as.data.frame(scale(World_Indicators_Scaling[,1:2]))
World_Indicators_GHS$Country<-World_Indicators_GH$Country

#############
#GDP & Health Exp/Capita


#Determine the optimal number of clusters using the silhouette method
fviz_nbclust(World_Indicators_GHS[,1:2], kmeans, method = "silhouette")

#Determine the optimal number of clusters using the elbow method
fviz_nbclust(World_Indicators_GHS[,1:2], kmeans, method = "wss")

#Find the kmeans of GDP and Infant Mortality
set.seed(55)
cluster.World_Indicators_GHS <- kmeans(World_Indicators_GHS[,1:2], 2, nstart = 20)
cluster.World_Indicators_GHS

#Graph of the GDP & Infectant Mortality Clustered
cluster.World_Indicators_GHS$cluster <- as.factor(
  cluster.World_Indicators_GHS$cluster)
colnames(World_Indicators_GHS)
WIGHS_Clustered <- plot_ly(World_Indicators_GHS, x = ~GDP, 
                           y = ~Health.Exp.Capita, 
                           color = ~cluster.World_Indicators_GHS$cluster)
WIGHS_Clustered <- WIGHS_Clustered %>% add_markers()
WIGHS_Clustered <- WIGHS_Clustered %>% layout(scene = 
                                                list(xaxis = list(title = 'X1'),
                                                     yaxis = list(title = 'X2')))
WIGHS_Clustered <- WIGHS_Clustered %>% layout(
  title = 'GDP vs. Health.Exp.Capita')

WIGHS_Clustered # plot 

#Hierarchical Clustering
#Using Hierarchical analysis to plot the clusters

#To find the distance between each clusters
dm_Gdp_healthcapita <- dist(World_Indicators_GHS[,1:2], method = 'euclidean')

# Hierarchical Clustering

set.seed(50)
hc_Gdp_healthcapita <- hclust(dm_Gdp_healthcapita,method = "average")

#Plotting Dendogram for hierarchical cluster

hc_Gdp_healthcapita_dendogram <- plot(hc_Gdp_healthcapita, main = "GDP Vs Health.Capita Dendogram")

#Using Cut-tree to identify the number of clusters

abline(h = 3.3, col = "blue")
#Since there is only 2 clusters we couldn't see the cut tree line
fit_Gdp_healthcapita <- cutree(hc_Gdp_healthcapita, k = 2)
#table(fit8, Data8$Class)
rect.hclust(hc_Gdp_healthcapita, k = 2, border = "red")

#Binding the data
World_Indicators_GHS <- cbind(World_Indicators_GHS, fit_Gdp_healthcapita)

#Plotting hierarchical clusters

World_IndicatorsGHS_hc_plot <- plot_ly(World_Indicators_GHS, x = ~GDP, y = ~Health.Exp.Capita, color = ~as.factor(World_Indicators_GHS$fit_Gdp_healthcapita))
World_IndicatorsGHS_hc_plot <- World_IndicatorsGHS_hc_plot %>% add_markers()
World_IndicatorsGHS_hc_plot <- World_IndicatorsGHS_hc_plot %>% layout(title = 'GDP Vs Infant.Mortality.Rate hierarchical Clustered')
World_IndicatorsGHS_hc_plot

#Self Organizing Maps - 

#Create the model
World_Indicators_GHS_SOM <- as.matrix(scale(World_Indicators_GHS[,1:2]))
World_Indicators_GHS_Grid <- somgrid(xdim = 5, ydim = 5, topo = "rectangular")
set.seed(55)
World_Indicators_GHS_SOM_Model <- som(X = World_Indicators_GHS_SOM, grid = World_Indicators_GHS_Grid)

#Plot the model
plot(World_Indicators_GHS_SOM_Model, type = "counts")

plot(World_Indicators_GHS_SOM_Model, type = "mapping")

plot(World_Indicators_GHS_SOM_Model, type = "codes")


#ADD INTERNAL VALIDATION 
#validation for 2 clusters
calinhara(World_Indicators_GHS[,1:2],cluster.World_Indicators_GHS$cluster,cn=2)
#validation for 3
cluster.World_Indicators_GHS_3 <- kmeans(World_Indicators_GHS[,1:2], 3, 
                                         nstart = 20)
cluster.World_Indicators_GHS_3
calinhara(World_Indicators_GHS[,1:2],cluster.World_Indicators_GHS_3$
            cluster,cn=3)

#validation for 4
cluster.World_Indicators_GHS_4 <- kmeans(World_Indicators_GHS[,1:2], 4, 
                                         nstart = 20)
cluster.World_Indicators_GHS_4
calinhara(World_Indicators_GHS[,1:2],cluster.World_Indicators_GHS_4$
            cluster,cn=4)
colnames(World_Indicators_GHS)

##################### Post CH Value Plot ##########################
WIGHS_Clustered_CH <- plot_ly(World_Indicators_GHS, x = ~GDP, 
                              y = ~Health.Exp.Capita, 
                              color = ~cluster.World_Indicators_GHS_4$cluster)
WIGHS_Clustered_CH <- WIGHS_Clustered_CH %>% add_markers()
WIGHS_Clustered_CH <- WIGHS_Clustered_CH %>% layout(scene = 
                                                      list(xaxis = list(title = 'X1'),
                                                           yaxis = list(title = 'X2')))
WIGHS_Clustered_CH <- WIGHS_Clustered_CH %>% layout(
  title = 'GDP vs. Health Capita')

WIGHS_Clustered_CH # plot 



#typeof(cluster.World_IndicatorsGI)

#add dunn index

unique(World_Indicators_GHS$Clust)
typeof(cluster.World_Indicators_GHS$cluster)
cluster.World_Indicators_GHS$cluster<-as.numeric(
  cluster.World_Indicators_GHS$cluster)
cluster.stats(dist(World_Indicators_GHS[,1:2],method ='euclidean'),
              cluster.World_Indicators_GHS$cluster)

cluster.World_Indicators_GHS$cluster<-as.numeric(
  cluster.World_Indicators_GHS$cluster)

x=dunn(clusters = cluster.World_Indicators_GHS$cluster, 
       Data = World_Indicators_GHS[,1:2])
cat("The value for dunn index is ,",x,"")

#Combine clusters with original table
World_Indicators_GHS$Clust <- cluster.World_Indicators_GHS$cluster



World_Indicators_GHS$Clust_CH<-cluster.World_Indicators_GHS_4$cluster

World_Indicators_GHS_Sub1<-subset(World_Indicators_GHS,Clust_CH==1)
World_Indicators_GHS_Sub2<-subset(World_Indicators_GHS,Clust_CH==2)
World_Indicators_GHS_Sub3<-subset(World_Indicators_GHS,Clust_CH==3)
World_Indicators_GHS_Sub4<-subset(World_Indicators_GHS,Clust_CH==4)
print(World_Indicators_GHS_Sub1,World_Indicators_GHS_Sub2,World_Indicators_GHS_Sub3,World_Indicators_GHS_Sub4)
```
Task 2: The above analysis caters to the analysis between gdp and health exp per capita. The expenditure varies country to country but is not heavily reliant on GDP. The health expenditure seems high even in low gdp and in some countries there is a very little spent on healthcare.
```{r}
#Analysis 3

#Data
#World_Indicators <- read.csv("/Users/kailash/Downloads/World Indicators (2).csv")

##########Scaling ######################

World_Indicators_GB<- World_Indicators %>% select(GDP, Business.Tax.Rate,
                                                  Country)
typeof(World_Indicators_GB$Business.Tax.Rate)
World_Indicators_GB<-na.omit(World_Indicators_GB)

World_Indicators_GBS<-as.data.frame(scale(World_Indicators_GB[,1:2]))
#DF<-as.data.frame(scale(World_Indicators_Scaling[,1:2]))
World_Indicators_GBS$Country<-World_Indicators_GB$Country

#GDP & Business Tax Rate


#Determine the optimal number of clusters using the silhouette method
fviz_nbclust(World_Indicators_GBS[,1:2], kmeans, method = "silhouette")

#Determine the optimal number of clusters using the elbow method
fviz_nbclust(World_Indicators_GBS[,1:2], kmeans, method = "wss")

#Find the kmeans of GDP and Infant Mortality
set.seed(55)
cluster.World_Indicators_GBS <- kmeans(World_Indicators_GBS[,1:2], 2, 
                                       nstart = 20)
cluster.World_Indicators_GBS

#Graph of the GDP & Business Tax Rate Clustered
cluster.World_Indicators_GBS$cluster <- as.factor(
  cluster.World_Indicators_GBS$cluster)
colnames(World_Indicators_GBS)
WIGBS_Clustered <- plot_ly(World_Indicators_GBS, x = ~GDP, 
                           y = ~Business.Tax.Rate, 
                           color = ~cluster.World_Indicators_GBS$cluster)
WIGBS_Clustered <- WIGBS_Clustered %>% add_markers()
WIGBS_Clustered <- WIGBS_Clustered %>% layout(scene = 
                                                list(xaxis = list(title = 'X1'),
                                                     yaxis = list(title = 'X2')))
WIGBS_Clustered <- WIGBS_Clustered %>% layout(
  title = 'GDP vs. Business Tax Rate')

WIGBS_Clustered # plot 

#Hierarchical Clustering
#Using Hierarchical analysis to plot the clusters

#To find the distance between each clusters
dm_Gdp_BTR <- dist(World_Indicators_GBS[,1:2], method = 'euclidean')

# Hierarchical Clustering 

set.seed(50)
hc_Gdp_BTR <- hclust(dm_Gdp_BTR,method = "average")

#Plotting Dendogram for hierarchical cluster

hc_Gdp_BTR_dendogram <- plot(hc_Gdp_BTR, main = "GDP Vs Buisness Tax Rate Dendogram")

#Using Cut-tree to identify the number of clusters

abline(h = 3.3, col = "blue")
fit_Gdp_BTR <- cutree(hc_Gdp_BTR, k = 2)
#table(fit8, Data8$Class)
rect.hclust(hc_Gdp_BTR, k = 2, border = "red")

#Binding the data
World_Indicators_GBS <- cbind(World_Indicators_GBS, fit_Gdp_BTR)

#Plotting hierarchical clusters

World_IndicatorsGBS_hc_plot <- plot_ly(World_Indicators_GBS, x = ~GDP, y = ~Business.Tax.Rate, color = ~as.factor(World_Indicators_GBS$fit_Gdp_BTR))
World_IndicatorsGBS_hc_plot <- World_IndicatorsGBS_hc_plot %>% add_markers()
World_IndicatorsGBS_hc_plot <- World_IndicatorsGBS_hc_plot %>% layout(title = 'GDP Vs Buisness Tax Rate hierarchical Clustered')
World_IndicatorsGBS_hc_plot

#Self Organizing Maps
World_IndicatorsGBS_SOM <- as.matrix(scale(World_Indicators_GBS[,1:2]))
World_Indicators_GBS_Grid <- somgrid(xdim = 5, ydim = 5, topo = "rectangular")
set.seed(55)
World_Indicators_GBS_SOM_Model <- som(X = World_IndicatorsGBS_SOM, 
                                      grid = World_Indicators_GBS_Grid)

#Plot the model
plot(World_Indicators_GBS_SOM_Model, type = "counts")

plot(World_Indicators_GBS_SOM_Model, type = "mapping")

plot(World_Indicators_GBS_SOM_Model, type = "codes")

#ADD INTERNAL VALIDATION 
#validation for 2 clusters
calinhara(World_Indicators_GBS[,1:2],cluster.World_Indicators_GBS$cluster,cn=2)
#validation for 3
cluster.World_Indicators_GBS_3 <- kmeans(World_Indicators_GBS[,1:2], 3, 
                                         nstart = 20)
cluster.World_Indicators_GBS_3
calinhara(World_Indicators_GBS[,1:2],cluster.World_Indicators_GBS_3$
            cluster,cn=3)

#validation for 4
cluster.World_Indicators_GBS_4 <- kmeans(World_Indicators_GBS[,1:2], 4, 
                                         nstart = 20)
cluster.World_Indicators_GBS_4
calinhara(World_Indicators_GBS[,1:2],cluster.World_Indicators_GBS_4$
            cluster,cn=4)
colnames(World_Indicators_GBS)
##################### Post CH Value Plot ##########################
colnames(World_Indicators_GBS)
WIGBS_Clustered_CH <- plot_ly(World_Indicators_GBS, x = ~GDP, 
                              y = ~Business.Tax.Rate, 
                              color = ~cluster.World_Indicators_GBS_4$cluster)
WIGBS_Clustered_CH <- WIGBS_Clustered_CH %>% add_markers()
WIGBS_Clustered_CH <- WIGBS_Clustered_CH %>% layout(scene = 
                                                      list(xaxis = list(title = 'X1'),
                                                           yaxis = list(title = 'X2')))
WIGBS_Clustered_CH <- WIGBS_Clustered_CH %>% layout(
  title = 'GDP vs. Business Tax Rate')

WIGBS_Clustered_CH # plot 


#Add dunn index

unique(World_Indicators_GBS$Clust)
typeof(cluster.World_Indicators_GBS$cluster)
cluster.World_Indicators_GBS$cluster<-as.numeric(
  cluster.World_Indicators_GBS$cluster)
cluster.stats(dist(World_Indicators_GBS[,1:2],method ='euclidean'),
              cluster.World_Indicators_GBS$cluster)

cluster.World_Indicators_GBS$cluster<-as.numeric(
  cluster.World_Indicators_GBS$cluster)

x=dunn(clusters = cluster.World_Indicators_GBS$cluster, 
       Data = World_Indicators_GBS[,1:2])
cat("The value for dunn index is ,",x,"")

#Combine clusters with original table
World_Indicators_GBS$Clust <- cluster.World_Indicators_GBS$cluster
World_Indicators_GBS$Clust_CH<-cluster.World_Indicators_GBS_4$cluster

#China, Japan, and USA represent cluster 1, all other countries (excluding NA) are represented by cluster 2.

#Part 3

colnames(World_Indicators_GBS)
World_Indicators_GBS_Sub1<-subset(World_Indicators_GBS,Clust_CH==1)
World_Indicators_GBS_Sub2<-subset(World_Indicators_GBS,Clust_CH==2)
World_Indicators_GBS_Sub3<-subset(World_Indicators_GBS,Clust_CH==3)
World_Indicators_GBS_Sub4<-subset(World_Indicators_GBS,Clust_CH==4)
print(World_Indicators_GBS_Sub1,World_Indicators_GBS_Sub2,World_Indicators_GBS_Sub3,World_Indicators_GBS_Sub4)



```
#GDP VS Business Tax Rate
Countries with low gdp have business tax rate varying towards both sides of the axis, Tax rates seem average or minimal in high gdp countries but they're higher in countries with low gdp.
```{r}
#Analysis 4

#Data
#World_Indicators <- read.csv("/Users/kailash/Downloads/World Indicators (2).csv")

#Male vs Female Life Expectancy vs. GDP

#Clean the data to only include required columns
World_Indicators_FM<- World_Indicators %>% select(GDP,13,14,21)
names(World_Indicators_FM)
typeof(World_Indicators_FM$Life.Expectancy.Female)
World_Indicators_FM<-na.omit(World_Indicators_FM)

World_IndicatorsFM<-as.data.frame(scale(World_Indicators_FM[,1:3]))
#DF<-as.data.frame(scale(World_Indicators_Scaling[,1:2]))
World_IndicatorsFM$Country<-World_Indicators_FM$Country

#Determine the optimal number of clusters using the silhouette method
fviz_nbclust(World_IndicatorsFM[,1:3], kmeans, method = "silhouette")

#Determine the optimal number of clusters using the elbow method
fviz_nbclust(World_IndicatorsFM[,1:3], kmeans, method = "wss")

#Find the kmeans of GDP and Infant Mortality
set.seed(55)
typeof(World_IndicatorsFM[1])
colnames(World_IndicatorsFM)
cluster.World_IndicatorsFM <- kmeans(World_IndicatorsFM[,1:3], 2, nstart = 20)
cluster.World_IndicatorsFM

#Graph of the GDP & Life of Male/Female
cluster.World_IndicatorsFM$cluster <- as.factor(cluster.World_IndicatorsFM$cluster)

FM_Clustered <- plot_ly(World_IndicatorsFM, x = ~Life.Expectancy.Female, y = ~`Life.Expectancy.Male`, z = ~GDP, color = ~cluster.World_IndicatorsFM$cluster)
FM_Clustered <- FM_Clustered %>% add_markers()
FM_Clustered <- FM_Clustered %>% layout(scene = list(xaxis = list(title = 'Female Life Expectancy'),
                                                     yaxis = list(title = 'Male Life Expectancy'),
                                                     zaxis = list(title = 'GDP')))
FM_Clustered <- FM_Clustered %>% layout(title = 'Female vs. Male Life Expectancy vs. GDP Clustered')

FM_Clustered

#Combine clusters with original table
World_IndicatorsFM$Clust <- cluster.World_IndicatorsFM$cluster

#Hierarchical Clustering
#Using Hierarchical analysis to plot the clusters

#To find the distance between each clusters
dm_Gdp_FM <- dist(World_IndicatorsFM[,1:2], method = 'euclidean')

# Hierarchical Clustering 

set.seed(50)
hc_Gdp_FM <- hclust(dm_Gdp_FM,method = "average")

#Plotting Dendogram for hierarchical cluster

hc_Gdp_FM_dendogram <- plot(hc_Gdp_FM, main = "GDP Vs Life Expectancy of both Male and Female Dendogram")

#Using Cut-tree to identify the number of clusters

abline(h = 3.3, col = "blue")
fit_Gdp_FM <- cutree(hc_Gdp_FM, k = 2)
#table(fit8, Data8$Class)
rect.hclust(hc_Gdp_FM, k = 2, border = "red")

#Binding the data
World_IndicatorsFM <- cbind(World_IndicatorsFM, fit_Gdp_FM)

#Plotting hierarchical clusters

World_IndicatorsFM_hc_plot <- plot_ly(World_IndicatorsFM, x = ~Life.Expectancy.Female, y = ~`Life.Expectancy.Male`, z = ~GDP, color = ~as.factor(World_IndicatorsFM$fit_Gdp_FM))
World_IndicatorsFM_hc_plot <- World_IndicatorsFM_hc_plot %>% add_markers()
World_IndicatorsFM_hc_plot <- World_IndicatorsFM_hc_plot %>% layout(title = 'GDP Vs Life Expectancy of both Male and Female')
World_IndicatorsFM_hc_plot

#In general, countries male and female life expectancies have a positive linear relationship and have a somewhat positive linear relationship with GDP.

#Self Organizing Maps - 

#Create the model
World_IndicatorsFM_SOM <- as.matrix(scale(World_IndicatorsFM[,1:3]))
World_IndicatorsFM_Grid <- somgrid(xdim = 5, ydim = 5, topo = "rectangular")
set.seed(55)
World_IndicatorsFM_SOM_Model <- som(X = World_IndicatorsFM_SOM, grid = World_IndicatorsFM_Grid)

#Plot the model
plot(World_IndicatorsFM_SOM_Model, type = "counts")

plot(World_IndicatorsFM_SOM_Model, type = "mapping")

plot(World_IndicatorsFM_SOM_Model, type = "codes")

#############
#ADD INTERNAL VALIDATION 
#validation for 2 clusters
calinhara(World_IndicatorsFM[,1:3],cluster.World_IndicatorsFM$cluster,cn=2)
#validation for 3
cluster.World_IndicatorsFM_3 <- kmeans(World_IndicatorsFM[,1:3], 3, nstart = 20)
cluster.World_IndicatorsFM_3
calinhara(World_IndicatorsFM[,1:3],cluster.World_IndicatorsFM_3$cluster,cn=3)

#validation for 4
cluster.World_IndicatorsFM_4 <- kmeans(World_IndicatorsFM[,1:3], 4, nstart = 20)
cluster.World_IndicatorsFM_4
calinhara(World_IndicatorsFM[,1:3],cluster.World_IndicatorsFM_4$cluster,cn=4)

##################### Post CH Value Plot ##########################
#Graph of the GDP & Infant Mortality Clustered
cluster.World_IndicatorsFM_4$cluster <- as.factor(cluster.World_IndicatorsFM_4$
                                                    cluster)

FM_Clustered_4 <- plot_ly(World_IndicatorsFM, x = ~Life.Expectancy.Female, 
                          y = ~Life.Expectancy.Male, 
                          z = ~GDP, color = ~cluster.World_IndicatorsFM_4$
                            cluster)
FM_Clustered_4 <- FM_Clustered_4 %>% add_markers()
FM_Clustered_4 <- FM_Clustered_4 %>% layout(scene = 
                                              list(xaxis = 
                                                     list(title = 'Female Life Expectancy'),
                                                     yaxis = list(title = 'Male Life Expectancy'),
                                                     zaxis = list(title = 'GDP')))
FM_Clustered_4 <- FM_Clustered_4 %>% 
  layout(title = 'Female vs. Male Life Expectancy vs. GDP Clustered')

FM_Clustered_4

#add dunn index
unique(World_IndicatorsFM$Clust)
typeof(cluster.World_IndicatorsFM$cluster)
cluster.World_IndicatorsFM$cluster<-as.numeric(
  cluster.World_IndicatorsFM$cluster)
cluster.stats(dist(World_IndicatorsFM[,1:3],method ='euclidean'),
              cluster.World_IndicatorsFM$cluster)

#cluster.World_IndicatorsGI$cluster<-as.numeric(cluster.World_IndicatorsGI$cluster)

x=dunn(clusters = cluster.World_IndicatorsFM$cluster, 
       Data = World_IndicatorsFM[,1:3])
cat("The value for dunn index is ,",x,"")

#Combine clusters with original table
World_IndicatorsFM$Clust <- cluster.World_IndicatorsFM$cluster
World_IndicatorsFM$Clust_CH<-cluster.World_IndicatorsFM_4$cluster
#China, Japan, and USA represent cluster 1, all other countries (excluding NA) are represented by cluster 2.
World_IndicatorsFM_Sub1<-subset(World_IndicatorsFM,Clust_CH==1)
World_IndicatorsFM_Sub2<-subset(World_IndicatorsFM,Clust_CH==2)
World_IndicatorsFM_Sub3<-subset(World_IndicatorsFM,Clust_CH==3)
World_IndicatorsFM_Sub4<-subset(World_IndicatorsFM,Clust_CH==4)
print(World_IndicatorsFM_Sub1,World_IndicatorsFM_Sub2,World_IndicatorsFM_Sub3,World_IndicatorsFM_Sub4)

```
Most countries with less female life expectancy are countries that have low GDP. Male life expectency is higher is countries where they have a higher GDP.

```{r Analysis 5}
World_Indicators_GIN<-World_Indicators %>% select(GDP,Internet.Usage,Country)
#World_Indicators_GI<- World_Indicators %>% select(GDP,Internet.Usage,                                                  Country)
typeof(World_Indicators_GIN$Internet.Usage)
World_Indicators_GIN<-na.omit(World_Indicators_GIN)

World_IndicatorsGIN<-as.data.frame(scale(World_Indicators_GIN[,1:2]))
#DF<-as.data.frame(scale(World_Indicators_Scaling[,1:2]))
World_IndicatorsGIN$Country<-World_Indicators_GIN$Country
library(fpc)
library(cluster)
library(factoextra)
#Determine the optimal number of clusters using the silhouette method

fviz_nbclust(World_IndicatorsGIN[,1:2], kmeans, method = "silhouette")

#Determine the optimal number of clusters using the elbow method
fviz_nbclust(World_IndicatorsGIN[,1:2], kmeans, method = "wss")

#Find the kmeans of GDP and Infant Mortality
set.seed(55)
cluster.World_IndicatorsGIN <- kmeans(World_IndicatorsGIN[,1:2], 2, nstart = 20)
cluster.World_IndicatorsGIN
set.seed(55)
cluster.World_IndicatorsGIN_sil <- kmeans(World_IndicatorsGIN[,1:2], 10, nstart = 20)
cluster.World_IndicatorsGIN_sil
#Graph of the GDP & Infant Mortality Clustered
cluster.World_IndicatorsGIN$cluster <- as.factor(cluster.World_IndicatorsGIN$cluster)
colnames(World_IndicatorsGIN)
WIGIN_Clustered <- plot_ly(World_IndicatorsGIN, x = ~GDP, y = ~Internet.Usage, color = ~cluster.World_IndicatorsGIN$cluster)
WIGIN_Clustered <- WIGIN_Clustered %>% add_markers()
WIGIN_Clustered <- WIGIN_Clustered %>% layout(scene = list(xaxis = list(title = 'X1'),
                                                           yaxis = list(title = 'X2')))
WIGIN_Clustered <- WIGIN_Clustered %>% layout(title = 'GDP vs. Internet Usage Clustered')

WIGIN_Clustered
############Silhoutte Graph#####################
cluster.World_IndicatorsGIN_sil$cluster <- as.factor(cluster.World_IndicatorsGIN_sil$cluster)

WIGIN_Clustered_Sil <- plot_ly(World_IndicatorsGIN, x = ~GDP, 
                               y = ~Internet.Usage, 
                               color = ~cluster.World_IndicatorsGIN_sil$cluster)
WIGIN_Clustered_Sil <- WIGIN_Clustered_Sil %>% add_markers()
WIGIN_Clustered_Sil <- WIGIN_Clustered_Sil %>% layout(
  scene = list(xaxis = list(title = 'X1'),
                                                           yaxis = list(title = 'X2')))
WIGIN_Clustered_Sil <- WIGIN_Clustered_Sil %>% 
  layout(title = 'GDP vs. Internet Usage Clustered')

WIGIN_Clustered_Sil

#Combine clusters with original table
World_IndicatorsGIN$Clust_1 <- cluster.World_IndicatorsGIN$cluster
World_IndicatorsGIN$Clust_Sil <- cluster.World_IndicatorsGIN_sil$cluster

#Similar to GDP and Infant Mortality... China, Japan, USA in cluster 1 and the rest in cluster 2.
#ADD INTERNAL VALIDATION 
#validation for 2 clusters
calinhara(World_IndicatorsGIN[,1:2],cluster.World_IndicatorsGIN$cluster,cn=2)
#validation for 3
cluster.World_IndicatorsGIN_3 <- kmeans(World_IndicatorsGIN[,1:2], 3, 
                                         nstart = 20)
cluster.World_IndicatorsGIN_3
calinhara(World_IndicatorsGIN[,1:2],cluster.World_IndicatorsGIN$cluster,cn=3)

#validation for 4
cluster.World_IndicatorsGIN_4 <- kmeans(World_IndicatorsGIN[,1:2], 4, 
                                         nstart = 20)
cluster.World_IndicatorsGIN_4

calinhara(World_IndicatorsGIN[,1:2],cluster.World_IndicatorsGIN_4$
            cluster,cn=4)
#colnames(World_Indicators_GHS)
#validation for 10
cluster.World_IndicatorsGIN_10 <- kmeans(World_IndicatorsGIN[,1:2], 10, 
                                        nstart = 20)
cluster.World_IndicatorsGIN_10

calinhara(World_IndicatorsGIN[,1:2],cluster.World_IndicatorsGIN_10$
            cluster,cn=10)

##################### Post CH Value Plot ##########################
WIINS_Clustered_CH <- plot_ly(World_IndicatorsGIN, x = ~GDP, 
                              y = ~Internet.Usage, 
                              color = ~cluster.World_IndicatorsGIN_10$cluster)
WIINS_Clustered_CH <- WIINS_Clustered_CH %>% add_markers()
WIINS_Clustered_CH <- WIINS_Clustered_CH %>% layout(scene = 
                                                      list(xaxis = list(title = 'X1'),
                                                           yaxis = list(title = 'X2')))
WIINS_Clustered_CH <- WIINS_Clustered_CH %>% layout(
  title = 'GDP vs. Internet Usage')

WIINS_Clustered_CH # plot 



#typeof(cluster.World_IndicatorsGI)

#add dunn index
#install.packages("philentropy")
#install.packages("clValid")

library(philentropy)
library(clValid)

unique(World_IndicatorsGIN$Clust)
typeof(cluster.World_IndicatorsGIN$cluster)
cluster.World_IndicatorsGIN$cluster<-as.numeric(
  cluster.World_Indicators_GHS$cluster)
library(clValid)
#cluster.stats(dist(World_IndicatorsGIN[,1:2]))
#cluster.stats(dist(World_IndicatorsGIN[,1:2],method ='euclidean'),
 #             cluster.World_IndicatorsGIN$cluster)

#cluster.World_IndicatorsGIN$cluster<-as.numeric cluster.World_IndicatorsGIN$cluster)

x=dunn(clusters = cluster.World_IndicatorsGIN$cluster, 
       Data = World_IndicatorsGIN[,1:2])
cat("dunn index is, ",x,"")
#Combine clusters with original table
#World_IndicatorsGIN$Clust <- cluster.World_IndicatorsGIN$cluster
World_IndicatorsGIN$Clust_CH<-cluster.World_IndicatorsGIN_4$cluster

#China, Japan, and USA represent cluster 1, all other countries (excluding NA) are represented by cluster 2.

#Part 3
#colnames(World_IndicatorsGIN)

#World_IndicatorsGIN_Sub1<-subset(World_IndicatorsGIN,Clust==1)
#World_IndicatorsGIN_Sub2<-subset(World_IndicatorsGIN,Clust==2)

#unique(World_IndicatorsGIN$Clust)

#Self Organizing Maps - 

#Create the model
World_IndicatorsGIN_SOM <- as.matrix(scale(World_IndicatorsGIN[,1:2]))
World_IndicatorsGIN_Grid <- somgrid(xdim = 5, ydim = 5, topo = "rectangular")
set.seed(55)
World_IndicatorsGIN_SOM_Model <- som(X = World_IndicatorsGIN_SOM, grid = World_IndicatorsGIN_Grid)

#Plot the model
plot(World_IndicatorsGIN_SOM_Model, type = "counts")

plot(World_IndicatorsGIN_SOM_Model, type = "mapping")

plot(World_IndicatorsGIN_SOM_Model, type = "codes")
```